HPC On-boarding

### 2016-01-19 hpc_101/paper.tex
adjustment of the partition image. added section describing the different slurm commands.

@@

\subsection{Cray CS300}
\begin{frame}
\frametitle{Cray CS300 -- History}
\begin{itemize}
		\item Brought online Fall 2014 -- Spring 2015
		\item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
		\item As of October 2015, more than 200 users have been granted access to the cluster
	\end{itemize}
\end{frame}

\subsubsection{Hardware}
\begin{frame}
	\frametitle{Cray CS300 -- Compute Nodes}
	\begin{itemize}
		\item 3,800 total cores -- \textbf{Spring upgrade} $\rightarrow$ 5,400$+$
		\item 178 standard nodes -- \textbf{Spring upgrade} $\rightarrow$ 270
			\begin{itemize}
				\item Two 10 core {\intel} processors (\emph{20 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx128$GB of useable RAM
			\end{itemize}
		\item 6 large memory nodes
			\begin{itemize}
				\item Four 10 core {\intel} processors (\emph{40 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx1$TB of useable RAM
			\end{itemize}			
		\item CentOS Linux
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Cray CS300 -- Storage}
	Two storage options are currently available on the Cray CS300.
	\begin{enumerate}
		\item {\lustre}
		\item ValueStorage
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{Cray CS300 -- Storage $\rightarrow$ {\lustre}}
	\begin{itemize}
		\item {\lustre} is a high performance parallel filesystem
		\item The Cray CS300 has $\approx582$TB of storage space
		\item Shared between all compute nodes and login nodes
		\item Primarily used as scratch space for jobs (Input and Output)
		\item User do not have a usage quota (soft or hard)
		\item Certain directories are subject to a 90 day purge policy
		\item \textbf{Data is not backed up!  Users are responsible for their own data}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Cray CS300 -- Storage $\rightarrow$ ValueStorage}
	\begin{itemize}
		\item $500$TB of scale out storage
		\item Currently only available for purchase by cluster users and only accessible via the login nodes
		\item ValueStorage will eventually be available for purchase by everyone
		\item ValueStorage owners will eventually be able to mount as a network drive (CIFS) on laptop, workstations, servers
		\item Purchased in 0.5TB increments
	\end{itemize}
	\btVFill

@@

\begin{frame}
	\frametitle{Cray CS300 -- Network}
	\begin{itemize}
		\item 40Gb Infiniband inter-connects (QDR)
		\begin{itemize}
			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
			\item Utilizes the \emph{fat tree network topology}
			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
		\end{itemize}
		\item 10Gb login node internet connectivity
		\begin{itemize}
			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
		\end{itemize}	
	\end{itemize}
\end{frame}

\subsubsection{Layout}
\begin{frame}
	\frametitle{Cray CS300 -- Layout}
	\includegraphics[width=0.95\textwidth]{images/high_level}
\end{frame}

@@

\subsection{Connecting to a cluster @ UH}
\begin{frame}
	\frametitle{Connecting to a cluster @ UH}
	\begin{description}[\setlength{\leftmargini}{0pt}]
	\item[] \begin{itemize}
		\item To connect to the cluster, we utilize a client which communicates using the Secure Shell (\emph{SSH}) protocol 
		\item Linux and MacOSX, typically have a SSH client already installed
		\item Windows typically does not come with an SSH client installed
		\item Windows 10 may come pre-installed with a SSH client, but it might not be stable
		\item Suggested SSH clients for Windows include:
		\begin{itemize}
			\item \href{http://www.hawaii.edu/askus/685}{SSH Secure Shell}~(SSH 3.2.9)
			\item \href{http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html}{Putty}
		\end{itemize}
	\end{itemize}
	\item[] ~\\
	\item[] \begin{itemize}
		\item The Cray CS300 has two login nodes:
		\begin{itemize}
			\item uhhpc1.its.hawaii.edu
			\item uhhpc2.its.hawaii.edu
		\end{itemize}
	\end{itemize}
	\end{description}
	{\large Let's attempt to login!}
\end{frame}

\subsubsection{Login to the Cray CS300}
\begin{frame}
\frametitle{Login to the Cray CS300}
	\begin{center}\textbf{Windows}\end{center}
	\hrule~\\
	\begin{itemize}
		\item If SSH 3.2.9 installed (Lab PCs have it installed)
		\item Open the start menu, and type ``SSH'' and you should see a program called ``SSH Secure File Terminal Client''
		\item Click ``Quick Connect'' and enter the following information:
			\begin{itemize}
			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu --OR-- uhhpc2.its.hawaii.edu
			\item[] \textbf{User Name:} Your UH User name e.g., user99
			\item[] \textbf{Port:} 22
			\end{itemize}
		\item Press ``Connect''
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Login to the Cray CS300}
	\begin{center}\textbf{Mac \& Linux}\end{center}
	\hrule~\\
	\begin{itemize}
		\item Open a terminal window
		\item Enter one of the following:
		\begin{itemize}
			\item ssh $<$UH User name$>$@uhhpc1.its.hawaii.edu
			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
			\item \textbf{Example:} ssh user99@uhhpc1.its.hawaii.edu
		\end{itemize}
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Globus}\footnotesize
If you already use Globus, the endpoints for the Cray CS300 are:
\begin{itemize}
\item hawaii\#UHHPC1
\item hawaii\#UHHPC1
\end{itemize}
~\\ 
For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the {\ci} website for more information:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

@@

	\end{block}
	\begin{itemize}
		\item We globally install popular/frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\tiny
			\item `module avail' -- list installed modules
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module unload $<$module name$>$' -- Remove named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
		\begin{itemize}\footnotesize
			\item Two ways to install software on the cluster
			\begin{itemize}\tiny
				\item Download compatible binaries
				\item Compile software from source
			\end{itemize}
		\end{itemize}
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

\begin{frame}
	\frametitle{Acquiring Software -- Binaries and/or Source}
	\begin{itemize}
		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the Cray CS300
		\begin{itemize}
			\item Binaries compiled as x86\_64 (64-bit) for CentOS 6.5 or RHEL6.5 should work
		\end{itemize}
		\item You may also download tar or zipped software/source code directly from the login nodes using tools like \textbf{wget} \& \textbf{curl}
		\item You may also clone source repositories using the correct software revision tool: \textbf{git}, \textbf{svn}, \textbf{hg}, \textbf{cvs}, etc.
	\end{itemize}
\end{frame}

@@

\subsection{Managing User Jobs}
\begin{frame}
\frametitle{Managing User Jobs}
User jobs all come in different shapes and sizes:  
\begin{itemize}
	\item Require multiple nodes working in concert towards a common goal (MPI)
	\item Require a single node, in which they use multiple threads work together (OpenMP, pthreads, TBB)
	\item Require a lot of cores to process a lot of data in an identical manner, yet none of the inputs have dependencies on another (HTC)
\end{itemize}
The Cray CS300 is capable of handle many different types of jobs, but in a multi-user environment with so many users, how do we turn chaos into organized chaos?
~\\
~\\
This looks like a job for a \textbf{\emph{job scheduler}}!
\end{frame}

\begin{frame}
	\frametitle{What is a Job Scheduler?}
	\begin{definition}{\footnotesize
		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing).
		}
	\end{definition}
	Basic features expected of a job scheduler include:
	\begin{itemize}\footnotesize
    \item Interfaces which help to define workflows and/or job dependencies
    \item Automatic submission of executions
    \item Interfaces to monitor the executions
    \item Priorities and/or queues to control the execution order of unrelated jobs
	\end{itemize}
	For the Cray CS300, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}.
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

\subsection{SLURM}
\begin{frame}
\frametitle{SLURM}
\begin{block}{How are jobs scheduled?}
User submitted jobs are assigned a priority using a fairshare algorithm.  Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job.   
\end{block}
\end{frame}

\begin{frame}
\frametitle{How to use SLURM}
SLURM has a series of commands, each of which allow users to interact with the job scheduler
\begin{itemize}
	\item sbatch -- 
	\item srun --
	\item srun.x11 --
	\item sacct --
	\item scontrol --
	\item scancel --
\end{itemize}  
~\\
Let us look at how the nodes are partitioned, as well as some examples of submitting jobs and create job scripts
\end{frame}

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The Cray CS300 currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
	\item More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
	\begin{itemize}\footnotesize
		\item[] \texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
		\item[] Example: \texttt{$[$login \ctilde$]$\$ scontrol show partition community.q}
	\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{SLURM Submission Script File (MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat mpi.slurm

@@

\begin{frame}[fragile]
\frametitle{SLURM Submission Script File (Non-MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat hello_world.slurm

@@

\begin{frame}[fragile]
\frametitle{SLURM Submission Script File (Job Array)}
\begin{semiverbatim}\tiny
[login lus]\$ cat job_array.slurm

>>>>

\newcommand{\craycs}{Cray~CS300}

@@

\subsection{{\craycs}}
\begin{frame}
\frametitle{{\craycs} -- History}
\begin{itemize}
		\item Brought online Fall 2014 -- Spring 2015
		\item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
		\item As of October 2015, more than 200 users have been granted access to the cluster
	\end{itemize}
\end{frame}

\subsubsection{Hardware}
\begin{frame}
	\frametitle{{\craycs}0 -- Compute Nodes}
	\begin{itemize}
		\item 3,800 total cores -- \textbf{Spring upgrade} $\rightarrow$ 5,400$+$
		\item 178 standard nodes -- \textbf{Spring upgrade} $\rightarrow$ 270
			\begin{itemize}
				\item Two 10 core {\intel} processors (\emph{20 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx128$GB of useable RAM
			\end{itemize}
		\item 6 large memory nodes
			\begin{itemize}
				\item Four 10 core {\intel} processors (\emph{40 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx1$TB of useable RAM
			\end{itemize}			
		\item CentOS Linux
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage}
	Two storage options are currently available on the {\craycs}
	\begin{enumerate}
		\item {\lustre}
		\item ValueStorage
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
	\begin{itemize}
		\item {\lustre} is a high performance parallel filesystem
		\item The {\craycs} has $\approx582$TB of storage space
		\item Shared between all compute nodes and login nodes
		\item Primarily used as scratch space for jobs (Input and Output)
		\item User do not have a usage quota (soft or hard)
		\item Certain directories are subject to a 90 day purge policy
		\item \textbf{Data is not backed up!  Users are responsible for their own data}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
	\begin{itemize}
		\item $500$TB of scale out storage
		\item Currently only available for purchase by cluster users and only accessible via the login nodes
		\item ValueStorage will eventually be available for purchase by everyone
		\item ValueStorage owners will eventually be able to mount as a network drive (CIFS) on laptop, workstations, servers
		\item Purchased in 0.5TB increments
	\end{itemize}
	\btVFill

@@

\begin{frame}
	\frametitle{{\craycs} -- Network}
	\begin{itemize}
		\item 40Gb Infiniband inter-connects (QDR)
		\begin{itemize}
			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
			\item Utilizes the \emph{fat tree network topology}
			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
		\end{itemize}
		\item 10Gb login node internet connectivity
		\begin{itemize}
			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
		\end{itemize}	
	\end{itemize}
\end{frame}

\subsubsection{Layout}
\begin{frame}
	\frametitle{{\craycs} -- Layout}
	\includegraphics[width=0.95\textwidth]{images/high_level}
\end{frame}

@@

\subsection{Connecting to a cluster @ UH}
\begin{frame}
	\frametitle{Connecting to a cluster @ UH}
	\begin{description}[\setlength{\leftmargini}{0pt}]
	\item[] \begin{itemize}
		\item To connect to the cluster, we utilize a client which communicates using the Secure Shell (\emph{SSH}) protocol 
		\item Linux and MacOSX, typically have a SSH client already installed
		\item Windows typically does not come with an SSH client installed
		\item Windows 10 may come pre-installed with a SSH client, but it might not be stable
		\item Suggested SSH clients for Windows include:
		\begin{itemize}
			\item \href{http://www.hawaii.edu/askus/685}{SSH Secure Shell}~(SSH 3.2.9)
			\item \href{http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html}{Putty}
		\end{itemize}
	\end{itemize}
	\item[] ~\\
	\item[] \begin{itemize}
		\item The {\craycs} has two login nodes:
		\begin{itemize}
			\item uhhpc1.its.hawaii.edu
			\item uhhpc2.its.hawaii.edu
		\end{itemize}
	\end{itemize}
	\end{description}
	{\large Let's attempt to login!}
\end{frame}

\subsubsection{Login to the {\craycs}}
\begin{frame}
\frametitle{Login to the {\craycs}}
	\begin{center}\textbf{Windows}\end{center}
	\hrule~\\
	\begin{itemize}
		\item If SSH 3.2.9 installed (Lab PCs have it installed)
		\item Open the start menu, and type ``SSH'' and you should see a program called ``SSH Secure File Terminal Client''
		\item Click ``Quick Connect'' and enter the following information:
			\begin{itemize}
			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu --OR-- uhhpc2.its.hawaii.edu
			\item[] \textbf{User Name:} Your UH User name e.g., user99
			\item[] \textbf{Port:} 22
			\end{itemize}
		\item Press ``Connect''
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Login to the {\craycs}}
	\begin{center}\textbf{Mac \& Linux}\end{center}
	\hrule~\\
	\begin{itemize}
		\item Open a terminal window
		\item Enter one of the following:
		\begin{itemize}
			\item ssh $<$UH User name$>$@uhhpc1.its.hawaii.edu
			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
			\item \textbf{Example:} ssh user99@uhhpc1.its.hawaii.edu
		\end{itemize}
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Globus}\footnotesize
If you already use Globus, the endpoints for the {\craycs} are:
\begin{itemize}
\item hawaii\#UHHPC1
\item hawaii\#UHHPC1
\end{itemize}
~\\ 
For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the {\ci} website for more information:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

@@

	\end{block}
	\begin{itemize}
		\item We globally install frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\footnotesize
			\item `module avail' -- list installed modules
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module unload $<$module name$>$' -- Remove named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
		%\begin{itemize}\footnotesize
			%\item Two ways to install software on the cluster
			%\begin{itemize}\tiny
				%\item Download compatible binaries
				%\item Compile software from source
			%\end{itemize}
		%\end{itemize}
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

\begin{frame}
	\frametitle{Acquiring Software -- Binaries and/or Source}
	\begin{itemize}
		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the {\craycs}
		\begin{itemize}
			\item Binaries compiled as x86\_64 (64-bit) for CentOS 6.5 or RHEL6.5 should work
		\end{itemize}
		\item You may also download tar or zipped software/source code directly from the login nodes using tools like \textbf{wget} \& \textbf{curl}
		\item You may also clone source repositories using the correct software revision tool: \textbf{git}, \textbf{svn}, \textbf{hg}, \textbf{cvs}, etc.
	\end{itemize}
\end{frame}

@@

\subsection{Managing User Jobs}
\begin{frame}
\frametitle{Managing User Jobs}
User jobs all come in different shapes and sizes:  
\begin{itemize}\footnotesize
	\item Require multiple nodes working in concert towards a common goal (MPI)
	\item Require a single node, in which they use multiple threads~\\work together (OpenMP, pthreads)
	\item Require a lot of cores to process a lot of data in an identical manner,~\\yet none of the inputs have dependencies on another (HTC)
\end{itemize}
\bigskip
The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we turn this chaos into organized chaos?
\btVFill
\begin{center}This looks like a job for a \textbf{\emph{job scheduler}}!\end{center}
\end{frame}

\begin{frame}
	\frametitle{What is a Job Scheduler?}
	\begin{definition}{\footnotesize
		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing)
		}
	\end{definition}
	Basic features expected of a job scheduler include:
	\begin{itemize}\footnotesize
    \item Interfaces which help to define workflows and/or job dependencies
    \item Automatic submission of executions
    \item Interfaces to monitor the executions
    \item Priorities and/or queues to control the execution order of unrelated jobs
	\end{itemize}
	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

\subsection{SLURM}
\begin{frame}
\frametitle{SLURM}
\begin{block}{How are jobs scheduled?}
User submitted jobs are assigned a priority using a fairshare algorithm.  Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job   
\end{block}
%The following slides that cover slurm will cover
%\begin{enumerate}
%\item Several useful SLURM commands
%\item Resource partitions
%\item example interactive jobs and batch job scripts
%\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{SLURM commands}\footnotesize
SLURM has a series of commands, each of which allow users to interact with the job scheduler
\begin{itemize}\tiny
	\item \emph{\textbf{srun}} --  Used to submit a job for execution or initiate job steps in real time
	\item \emph{\textbf{srun.x11}} -- Used to submit an interactive job with x11 support
	\item \emph{\textbf{sbatch}} -- Used to submit a job script for later execution. The script could contain one or more srun commands
	\item \emph{\textbf{scancel}} -- Used to cancel a pending or running job or job step. It can also be used to send an arbitrary signal to all processes associated with a running job or job step
	\item \emph{\textbf{sinfo}} -- Reports the state of partitions and nodes managed by Slurm. It has a wide variety of filtering, sorting, and formatting options
	\item \emph{\textbf{squeue}} -- Reports the state of jobs or job steps
	\item \emph{\textbf{sacct}} -- Used to report job or job step accounting information about active or completed jobs
	\item \emph{\textbf{scontrol}} -- The administrative tool used to view and/or modify Slurm state. Note that many scontrol commands can only be executed as user root
\end{itemize}
\begin{itemize}
	\item[--] Examples usage of the SLURM commands can be seen on schedmd's \href{http://slurm.schedmd.com/quickstart.html}{quickstart}
	\item[--] Each command should have a `man' page, or displays help when the -h flag is used
\end{itemize}
\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The {\craycs} currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
	\item More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
	\begin{itemize}\footnotesize
		\item[] \texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
		\item[] Example: \texttt{$[$login \ctilde$]$\$ scontrol show partition community.q}
	\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat mpi.slurm

@@

\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Non-MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat hello_world.slurm

@@

\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
\begin{semiverbatim}\tiny
[login lus]\$ cat job_array.slurm
